---
title: "sealevelmonitor"
author: "Willem Stolte, Fedor Baart"
date: "2024-03-12"
output:
  html_notebook:
    code_folding: hide
params:
  wind_or_surge_type: GTSM
  station: [Delfzijl, Netherlands (without Delfzijl)]
  modeltype: [linear, broken_linear]
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  comment=FALSE,
  message=FALSE
)

config <- RcppTOML::parseToml("_common/configuration.TOML")
source("_common/functions.R")
epoch = config$constants$epoch
mainstations_df <- readMainStationInfo()

```

## Read data

Monthly and yearly mean sea level is loaded from [PSMSL](https://psmsl.org/data/obtaining/) for the main tide gauge stations in the Netherlands. 

!!!!! Onderstaand vervangen door inlezen van origineel gedownloade data 
- psmsl
- gtsm
- wind heranalyses


```{r readData, comment=F, message=F}

df <- readSeaLevelData(config$constants$dataUrl) %>%
  addPreviousYearHeight() %>%
  addSurgeAnomaly() %>%
  addBreakPoints() %>%
  selectCols() %>%
  filter(
    station %in% params$station
         )
```



```{r}
df %>%
  # left_join(mainstations_df, by = c(station = "name")) %>%
  dplyr::group_by(station) %>%
  dplyr::summarise(start = min(year), end = max(year))
```



To do

-   read NCEP and other wind reanalysis data
-   make possibility to loop over all wind/surge correction options


```{r nest-per-station}
byStation <- df %>%
  dplyr::group_by(station) %>%
  tidyr::nest() %>%
  dplyr::ungroup()
```


```{r check-workflow}
wind_or_surge_type = params$wind_or_surge_type
check_workflow_wind(wind_or_surge_type = wind_or_surge_type)
```


```{r}

selectedmodel <- config$constants$modeltypes[config$constants$modeltypes %in% params$modeltype]
# model_functions_list <- map(paste(selectedmodel, "model", sep = "_"), get)

models <- byStation %>%
  # grid extended with model types
  expand_grid(modeltype = selectedmodel) %>%
  # add functions for model calculation
  mutate(modelfunctions = map(
    paste(modeltype, "model", sep = "_"), 
    get)
  ) %>%
  # add models based on data and functions
  mutate(model = pmap(
    list(
      data,
      modelfunctions
    ),
    \(.d, .f) .f(.d)
  )) %>%
  mutate(
    resids = map2(data, model, add_residuals)
  ) %>%
  mutate(
    preds = map2(data, model, add_predictions)
  ) %>%
  # add some model characteristics and skill terms
  mutate(
    glance = map(model, broom::glance),
    rsq    = glance %>% map_dbl("r.squared"),
    adj.rsq = glance %>% map_dbl("adj.r.squared"),
    AIC    = glance %>% map_dbl("AIC"),
    tidy   = map(model, broom::tidy),
    augment = map(model, broom::augment),
    equation = map(model, function(x) equatiomatic::extract_eq(x))
  )

```

## Autocorrelation

Autocorrelation

```{r, fig.width=6, fig.height=3}
models %>%
  mutate(
    ACF = map(augment, function(x) fortify(acf(x$.resid, plot = F)))
  ) %>%
  unnest(ACF) %>%
  ggplot(aes(Lag, ACF)) +
  geom_line() +
  geom_vline(xintercept = 8.9, linetype = 3) +
  geom_vline(xintercept = 18.6, linetype = 3) +
  facet_grid(station ~ modeltype) +
  theme(strip.text.y = element_text(angle = 0))

```
## Heteroskedasticity

### Residuals

```{r, fig.width=8, fig.height=4}
models %>%
  # filter(!grepl("Netherlands", station)) %>%
  # filter(station == "IJmuiden") %>%
  unnest(c(data, resids), names_sep = "_") %>% #str(max.level = 2)
ggplot(aes(x = resids_resid)) +
  geom_histogram(bins = 20, aes()) +
  facet_grid(modeltype ~ station) #+

```



```{r}
models %>%
  # filter(!grepl("Netherlands", station)) %>%
  # filter(station == "IJmuiden") %>%
  unnest(c(data, resids), names_sep = "_") %>% #str(max.level = 2)
ggplot(aes(data_year, resids_resid)) +
  geom_point(alpha = 0.4) +
  facet_grid(modeltype ~ station) #+
```


```{r}
models %>%
  # filter(!grepl("Netherlands", station)) %>%
  # filter(station == "IJmuiden") %>%
  unnest(c(data, resids), names_sep = "_") %>% #str(max.level = 2)
ggplot(aes(`resids_surge anomaly`, resids_resid)) +
  geom_point(alpha = 0.4) +
  facet_grid(modeltype ~ station) #+
```


## Sea level rise

```{r, fig.width=12, fig.height=6}
models %>%
  # filter(grepl("Netherlands", station)) %>%
  # filter(station == "IJmuiden") %>%
  # filter(modeltype == "linear") %>%
  unnest(c(data, preds), names_sep = "_") %>% #str(max.level = 2)
  # mutate(
  #   xx = min(data_year + epoch), 
  #   yy = max(`data_height - surge anomaly`)
  #   ) %>%
ggplot(aes(data_year, data_height - `data_surge anomaly`)) +
  geom_point(alpha = 0.4) +
  geom_line(aes(
    x = preds_year, 
    y = preds_pred - (`preds_surge anomaly`), 
    color = modeltype
  ),
  linewidth = 1) +
  # geom_line(
  #   aes(`augment_I(year - epoch)` + epoch, 
  #       augment_.fitted,
  #       color = modeltype), size = 1) +
  # geom_text(aes(x = xx, y = yy, hjust = "inward", vjust = "inward", label = equation), parse = TRUE, size = 2.5) +
  facet_wrap(~ station) #+
  # coord_cartesian(
  #   xlim = c(1980, NA),
  #   ylim = c(-100, NA)
  #   )

```


## Parameters



```{r, fig.height=10}

models %>% 
  select(station, modeltype, tidy) %>% 
  unnest(tidy) %>% 
  kableExtra::kable(caption = "Coefficients for all models and stations.") %>%
  kableExtra::scroll_box(height = "500px")
```


